---
title: "Machine Learning Assignement"
author: "Andrea Alberto"
date: "19 December 2017"
output:
  html_document: default
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
	echo=TRUE, warning=FALSE, message=TRUE,
	cache=TRUE)
library(ggplot2)	
library(plyr)
library(caret)
```

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

```{r}
# training <- read.csv(url('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv',na.strings = c("NA","NaN","","#DIV/0!","")))
# testing <- read.csv(url('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv',na.strings = c("NA","NaN","","#DIV/0!",""))

training <- read.csv("D:\\Personal\\Projects\\My\\DataScienceToolboxCourse\\prjs\\pml-training.csv",header = TRUE,na.strings = c("NA","NaN","","#DIV/0!",""))
testing <- read.csv("D:\\Personal\\Projects\\My\\DataScienceToolboxCourse\\prjs\\pml-testing.csv",header = TRUE,na.strings = c("NA","NaN","","#DIV/0!",""))
```

# Data Processing
We remove the columns we won't use
```{r}
badColumns1 <- c('X','raw_timestamp_part_1','raw_timestamp_part_2','cvtd_timestamp','new_window')
training <- training[,!(names(training) %in% badColumns1)]
```

Check the number of na per each predictor
```{r}
naVars <- sapply(training, function(x) sum(is.na(x))) > (.9 * nrow(training))
sum(naVars)
```

Out of 160 variables 100 are full of NAs. We won't use them.

Keep only the variable with less the 90% of NAs
```{r}
namesNotNaVars <- names(naVars[!naVars])
training <- training[,namesNotNaVars]
```

# Model fit and selection
Let's split the training data in training and test to forecast the out-of-bag accuracy
```{r}
inTrain <- createDataPartition(training$classe, p = 0.6)[[1]]
test <- training[-inTrain,]
training <- training[ inTrain,]
```

Fit a randomForest model which has included cross-validation
```{r}
set.seed(1234)
rfFit <- train(classe ~ . , data=training, method='rf', importance = T, trControl = trainControl(method = "cv", number = 3))
rfFit
```
Accuracy 0.9898951

Fit a boosting model which has included cross-validation
```{r boostTrain}
set.seed(1234)
garbage <- capture.output(
    boostFit <- train(classe ~ . , data=training, method='gbm', trControl = trainControl(method = "cv", number = 3))
)
boostFit
```
Accuracy 0.9841200


# Testing
Let's get an estime of out-of-bag accuracy

Random Forest model
```{r}
rfTestClasse <- predict(rfFit,test)
confusionMatrix(test$classe,rfTestClasse)
```
Accuracy : 0.9969

Boost model
```{r boostPredict}
boostTestClasse <- predict(boostFit,test)
confusionMatrix(test$classe,boostTestClasse)
```
Accuracy : 0.9892


# Validation
We can now predict the 'classe' of the validation data set
```{r}
validation <- testing[,!(names(testing) %in% badColumns1)]
validation <- validation[,namesNotNaVars[!(namesNotNaVars %in% c('classe'))]]

rfValidationClasse <- predict(rfFit,validation)
rfValidationClasse
```